\sysfull

### 2018-08-04 paper/asplos19/hwswsetup.tex
section 7 revised

\subsection{Intermittently-powered Platform and Experimental Setup}
\label{sec:results_hardware_software}

We used two platforms to evaluate \sys: (i) an \emph{intermittently-powered}
energy-harvesting device (WISP\,5.1~\cite{wisp5,wisp}), and (ii) a \emph{continuously-powered} evaluation
board (MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}).
%
Every benchmark listed in the next section was run repeatedly on
each platform for a few minutes to ensure capturing a diverse power 
trace and produce a high number of complete runs of all applications.
The results were averaged across the duration of the experiment.

% Experiments on the intermittently-powered platform used the
% WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
%
The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
configured to 1\,MHz clock speed.
%
We powered the WISP using an RF signal generator emitting  a 20\,dBm sinusoidal wave at 915\,MHz.
The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
antenna~\cite{atlas2015}.
%
The antenna was oriented towards and in parallel with the WISP antenna, and
no objects obstructed the path.
%
We affixed the WISP with a paper harness at the edge of a table at a height of 10\,cm.
%
For distance-controlled experiments we positioned WISP at $d=\{15, 30,
50\}$\,cm from the exciter antenna and 60\,cm was used for the task downscaling
experiments.
%
% We chose relatively shorter distances than in a deployment scenario to reduce
% charge time and run experiments faster, after having confirmed ourselves a
% result also reported in prior work~\cite{alpaca} that the amount of energy
% available in one charge cycle---the sum of stored energy and incoming energy
% during execution---is the same at distances beyond 50\,cm where incoming energy
% becomes negligible.
%
To obtain execution time, the software toggled GPIO pins at sections of the code
under profile, and the Saleae~\cite{saleae} logic analyzer measured
the interval between edges in the signal. When the board is continously powered the 
execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
%

\subsection{Software Benchmarks}
\label{sec:software_benchmarks}

@@

We evaluated \sys using six benchmarks summarized in
Table~\ref{table:benchmark_table}. %
Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) are from prior
work~\cite{chain,alpaca} and three benchmarks (\textbf{dijkstra},
\textbf{fft}, \textbf{sort}) are new implementations of applications
from the MiBench suite~\cite{mibench,hicks_mibench2_2016}. The
\textbf{fft} implementation is based on the Accelerated FFT library from
TI~\cite{ti_dsp}. All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
from toolchain version 5.01.01.00 with \texttt{-O1} as optimization flag.
%
The source code for all benchmarks will be released at \cite{coala_website}.

%Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code is accessible~\cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence. Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{dijkstra}: Dijkstra algorithm executed continously on a predefined adjecency matrix; \textbf{fft}: continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers. \todo{justify the selection of these benchmarks}

\textbf{Comparison with Alpaca using the GCC compiler.} We compare
\sys against Alpaca~\cite{alpaca}, a state-of-the-art task-based system for
intermittent computing. For a fair comparison the task decomposition is the
same in \sys and Alpaca implementations of the benchmarks. We compare
both systems using GCC.

Since Alpaca's compiler pass is implemented only for LLVM, we could not use
that implementation to instrument the benchmarks and compile them with GCC.
Instead, we \emph{manually} performed the instrumentation done by the compiler
pass.  We performed the instrumentation at the source level, adding C
statements to the program, instead of at the level of binary
intermediate-representation (LLVM IR) as the compiler pass would. The result is
equivalent from the perspective of overheads. The instrumentation consists of
identifying write-after-read dependencies and adding code and memory
allocations to make a private copy of the affected variables. For array
variables, instrumentation also includes adding code for 
dynamically tracking the array access.

Manual instrumentation of Alpaca code is not a general approach, but only
served as a practical alternative to porting the compiler pass from LLVM to
GCC.
%
Our experiments show that manually-instrumented Alpaca code compiled with GCC
is about 30\%-40\% faster than Alpaca code instrumented automatically by the
Alpaca LLVM compiler pass and compiled with LLVM. This performance difference
is due to the early experimental state of the MSP430 backend for
LLVM~\cite{baghsorkhi_cgo_2018}.
%
By compiling both systems with the same compiler (GCC), we ensure that our
comparison in Section~\ref{sec:evaluation} is fair.\vspace{2cm}

>>>>

\subsection{Experimental Setup}
\label{sec:results_hardware_software}

We used two platforms to evaluate \sys: (i) an \emph{intermittently-powered}
energy-harvesting device (WISP\,5.1~\cite{wisp5,wisp}), and (ii) a \emph{continuously-powered} evaluation
board (MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}).
%
Every benchmark listed in Section~\ref{sec:evaluation} was run repeatedly on
each platform for a few minutes to ensure capturing a diverse power 
trace and produce a high number of complete runs of all applications.
The results were averaged across the duration of the experiment.

The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
configured to 1\,MHz clock speed.
%
We powered the WISP using an RF signal generator emitting  a 20\,dBm sinusoidal wave at 915\,MHz.
The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
antenna~\cite{atlas2015}.
%
The antenna was oriented towards and in parallel with the WISP antenna, and
no objects obstructed the path.
%
We affixed the WISP with a paper harness at the edge of a table at a height, from the table suface, of 10\,cm.
%
For distance-controlled experiments we positioned WISP at $d=\{15, 30,
50\}$\,cm from the exciter antenna and 60\,cm was used for the task downscaling
experiments.
%
% We chose relatively shorter distances than in a deployment scenario to reduce
% charge time and run experiments faster, after having confirmed ourselves a
% result also reported in prior work~\cite{alpaca} that the amount of energy
% available in one charge cycle---the sum of stored energy and incoming energy
% during execution---is the same at distances beyond 50\,cm where incoming energy
% becomes negligible.
%
To obtain execution time, the software toggled GPIO pins at sections of the code
under profile, and the Saleae~\cite{saleae} logic analyzer measured
the interval between edges in the signal. When the board is continously powered the 
execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
%
\subsection{Software Benchmarks}
\label{sec:software_benchmarks}

@@

We evaluated \sys using six benchmarks that are oftenly used in embedded systems (summarized in
Table~\ref{table:benchmark_table}). All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 toolchain with \texttt{-O1} as optimization flag. The source code for all benchmarks will be released at \cite{coala_website}.

\textbf{Comparison with Alpaca using the GCC compiler.} We compare
\sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for
intermittent computing. For a fair comparison the task decomposition of the benchmarks are ensured to be the
same for both systems.

Since Alpaca's compiler pass is implemented only for LLVM, we could not use
that implementation to instrument the benchmarks and compile them with GCC.
Instead, we \emph{manually} performed the instrumentation done by the compiler
pass. The instrumentation consists of identifying write-after-read dependencies 
and adding code and memory allocations to make a private copy of the affected variables. 
For array variables, instrumentation also includes adding code for 
dynamically tracking the array access. Manual instrumentation of Alpaca code is not a general approach, 
but only served as a practical alternative to porting the compiler pass from LLVM to GCC.
%
Our experiments show that manually-instrumented Alpaca code compiled with GCC
is about 30\%-40\% faster than Alpaca code instrumented automatically by the
Alpaca LLVM compiler pass and compiled with LLVM. This performance difference
is due to the early experimental state of the MSP430 backend for
LLVM~\cite{baghsorkhi_cgo_2018}.
%
By compiling both systems with the same compiler (GCC), we ensure that our
comparison in Section~\ref{sec:evaluation} is fair.
